---
title: "Visualization using plotly and Leaflet packages for Corona Disease"
author: "Ozancan Özdemir"
date: "02 02 2020"
output: html_document
---

In this tutorial, I will try to show to create an interactive chart by using ```plotly``` and ```leaflet```  function by using Corona dataset taken from the Kaggle. (https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset) 
The source of the dataset: **Johns Hopkins University**




```{r,message=FALSE,warning=FALSE}
#install.packages("plotly")
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
```

The dataset shows the following information related to nCoV virus, known as Corona.

+ County that the virus seen. 

+ Provinces that virus seen.

+ Total number of patients having the virus.

+ Number of deaths caused by virus.

+ Total number of recovered patiens. 


Before reading the dataset, I divided dataset into two parts. I extracted the date column manually, then made a correction to simplify the analysis. 

```{r}
corona=read.csv("ncov_2019.csv",header=T,sep=";")
date=read.table("date.txt",header=T)
corona=data.frame(corona,date)
head(corona)
```

To draw the plot, I changed the class of the date object as date. 

```{r}
corona$Date=as.Date(corona$Date)
```

**Total Number of People having the disease between 22.01.2020-31.01.2020**

I calculated the sum of the confirmed number of patients with respect to dates between 22.01.2020-31.01.2020.

```{r}
sum_data=group_by(corona, Date) %>% summarize(sum_confirm = sum(Confirmed))
sum_data
```

```{r}
p <-sum_data %>%
  ggplot( aes(x=Date, y=sum_confirm)) +
  geom_area(fill="#69b3a2", alpha=0.5) +
  geom_line(color="#69b3a2") +
  ylab("Total Number of Patients")+
  theme_ipsum()

# Turn it interactive with ggplotly
p <- ggplotly(p)
p

```

It is seen that the spread of the disease approximately shows linearly increasing trend over time. 


**Total Number of Deaths caused by the disease between 22.01.2020-31.01.2020**

The total number of deaths caused by the disease is also visualized below. 

```{r}
death_data=group_by(corona, Date) %>% summarize(sum_death = sum(Deaths))
death_data
```

```{r}
p1 <-death_data %>%
  ggplot( aes(x=Date, y=sum_death)) +
  geom_area(fill="#CC0000", alpha=0.5) +
  geom_line(color="#CC0000") +
  ylab("Total Number of Deaths caused by Corona") +
  theme_ipsum()

# Turn it interactive with ggplotly
p1 <- ggplotly(p1)
p1

```

As seen that, the number of death patients shows similar behavior like spread of the disease. The number of deaths unfortunately shows increasing trend over time. 



**Total Number of Recovered Patients  between 22.01.2020-31.01.2020**


Now, the total number of recovered patients is also visualized below. 


```{r}
rec_data=group_by(corona, Date) %>% summarize(sum_rec = sum(Recovered))
rec_data
```

```{r}
p2 <-rec_data %>%
  ggplot( aes(x=Date, y=sum_rec)) +
  geom_area(fill="#D55E00", alpha=0.5) +
  geom_line(color="#D55E00") +
  ylab("Total Number of Deaths caused by Corona") +
  theme_ipsum()

# Turn it interactive with ggplotly
p2 <- ggplotly(p2)
p2

```


**Drawing a map using Leaflet**

First of all, we created the dataset showing total number of disease for each country. 

```{r}
country_data=group_by(corona, Country) %>% summarize(total= sum(Confirmed))
country_data
```

In order to draw a map using leaflet, we need to have lattiude and longitude of the countries. The following data set provides these information for the countries listed above. 

```{r}
lattitude_longitude=read.table("lattitude_longitude.txt",sep = "\t",header=T)
lattitude_longitude
```

Then, merge the data sets. 

```{r}
leaflet_data=data.frame(country_data,lattitude_longitude)
leaflet_data
```



```{r,message=FALSE,warning=FALSE}
library(leaflet)
# Center points for the map
center_lon <- median(leaflet_data$longitude, na.rm = TRUE)
center_lat <- median(leaflet_data$latitude, na.rm = TRUE)

leaflet(leaflet_data) %>%
  addProviderTiles("Esri") %>%
  addCircles(~leaflet_data$longitude, ~leaflet_data$latitude, weight =~leaflet_data$total^(1/2), popup = ~leaflet_data$name, color = "red") %>%
  setView(lng = center_lon, lat = center_lat, zoom = 2)

```

You can find the codes and dataset from my GitHub page. (https://github.com/ozancanozdemir)
